---
title: "Exploratory Analysis: Voter Disparities Among Marginalized Communities"
author: "Audrey Chen, Selina Dinh, Hadar Dolev"
date: "11/12/2020"
output: html_document
---
#Exploratory Analysis

```{r setup, include = FALSE}
library("knitr")
library("stringr")
library("tidyverse")
library("ggplot2")
source("scripts/chartOne.R")
source("scripts/chartTwo.R")
source("scripts/chartThree.R")
source("scripts/summaryInformation.R")
source("scripts/summaryInformationTable.R")
```

## Introduction
For the project, our area of focus is centered on analyzing the voting trends and 
patterns in the most recent elections. Specifically, we focused on the voting 
in the 2020 election between Joe Biden and Donald Trump and looked at the disparities
in voting trends between each state and county. Our data set also shows us the 
demographics of the citizens in those counties, so we are also able to notice
if there are any trends in voting based on gender, race, and any intersectionality
between the demographics. We are interested in this because we all feel that issues
and policies in politics are heavily swayed by gender and race. Due to that, analyzing 
the opinions of these groups can help us identify the differences in policies 
and appeal of each candidate. These issues are especially prevalent today as 
with the most recent election, most ideas were very polarized and driven by 
prejudices. With that, our group aims to dive more in-depth into this study
to observe where in the United States each political opinion alligns in order 
to compare that with the demographics in each location. With this information,
we can better understand the effects of race and gender on voting trends today.
Our data set is taken from a csv file with an accumulation of voting data.
and displays the voting information separated by location and the demographic of 
each location. This accumulation of information allows us to analyze any 
disparities or trends in voting data within our target marginalized communities.

# Insert summary paragpraph that includes information calculated by your
# summary information function 
As we want to investigate the relationship between Individuals who voted and the election 
results we wanted to find out the state/federal district with the most votes for both parties in thre 2020 election. 
``r max_percent_joe`` was the district with the most votes for Biden when the state with the most votes for Trump  was 
``r max_percent_trump``. We then looked for the state that has the highest population of Black residents which happens to be 
``r max_race_pop_black``. We found this to be surprising since ``r max_race_pop_black``
was a red state in the 2020 election. We decided to challenge this discovery and calculate the voter turnout
in the state which ended up being ``r voter_turnout_ms``%. This rose a lot of hypotheses between our team such as voter suppression. 
We then decided to check the state with the highest white population and we discovered it was
``r max_race_pop_white``, which was in fact, a red state as well. 

# Render table and describe pertinent observations found in it 
```{r summaryInformationTable, echo = FALSE, message =  FALSE}
```
This table reveals many insights to voter behavior. It shows total  individuals voted, 
total population, and total number registered from each race and gender. We can gather 
a lot of information on who is more likely to vote and what other affects their access 
to voting. 

# Describe the purpose of each chart, render each chart, and describe pertinent
# information found in each chart 

##chart 1
```{r bar_2016, echo = FALSE, message =  FALSE}
```
```{r bar_2020, echo = FALSE, message =  FALSE}
```  
For the first chart, we decided to create two separate bar charts to showcase 
the number of votes per state in the 2016 presidential election, `total_votes16`, versus the number of votes per states in the 2020 presidential election, `total_votes20`. By doing this, we are able to observe which states had higher voter turnout, had lower voter turnout, as well as the difference between the two 
different elections. In addition, with this information, we can apply it to what we know about voter disparities and analyze if there are any correlation between that information and the evident numeric data. 

##chart 2
```{r race_prop, echo = FALSE, message =  FALSE}
```
For the second chart, our group focused on proportions. Specifically, which
proportions of state populations do the racial groups of Hispanic, White, Black,
Native, Asian, Pacific amount to. To do that, we utilized a stacked bar chart, as 
we felt that it would be the most fitting visual to showcase this analysis.
Observing this information allowed us to see which racial groups holds 
a majority position in which state, and whether or not this demographic trends
leverages any affects for voter turnout and preference of a certain candidate.

##chart 3
```{r votes_2020, echo = FALSE, message =  FALSE}
```
For the third chart, a scatterplot was created to compare the votes for Donald 
Trump versus for Joe Biden in the 10 out of 13 considered swing states for the 
2020 election. By analyzing this, we wanted to see the trends and difference in votes between the two election's candidate for the specified swing states, as well as the 
amount of votes that ultimately lead to the winner of that state's electoral votes. 
