---
title: "Exploratory Analysis: Voter Disparities Among Marginalized Communities"
author: "Audrey Chen, Selina Dinh, Hadar Dolev"
date: "11/12/2020"
output: html_document
---
#Exploratory Analysis

```{r setup, include = FALSE}
library("knitr")
library("stringr")
library("tidyverse")
library("ggplot2")
source("scripts/chartOne.R")
source("scripts/chartTwo.R")
source("scripts/chartThree.R")
source("scripts/summaryInformation.R")
source("scripts/summaryInformationTable.R")
```

## Introduction
For the project, our area of focus is centered analyzing the voting trends and 
patterns in the most recent elections. Specifically, we focused on the voting 
in the 2020 election between Joe Biden and Donald Trump and looked at the disparities
in voting trends between each state and county. Our data set also shows us the 
demographics of the citizens in those counties, so we are also able to notice
if there are any trends in voting based on gender, race, and any intersectionality
between the demographics. We are interested in this, because we all feel that issues
and policies in politics are heavily swayed by gender and race and so analyzing 
the opinions of these groups can help us identify the differences in policies 
and appeal of each candidate. These issues are especially prevalent today as 
with the most recnet election, most ideas were very polarized and driven by 
prejudices. With that, our group aims to dive more in-depth into this study
to observe where in the United States each political opinion alligns in order 
to compare that with the demographics in each location. With this information,
we can better understand the effects of race and gender on voting trends today.
Our data set is taken from a csv file with an accumulation of voting data.
and displays the voting information separated by location and the demographic of 
each location. This accumulation of information allows us to analyze any 
disparities or trends in voting data within our target marginalized communities.

# Insert summary paragpraph that includes information calculated by your
# summary information function 
As we want to investigate the relationship between Individuals who voted and the election 
results we wanted to find out the state/federal district with the most votes for both parties in thre 2020 election. 
``r max_percent_joe`` was the district with the most votes for Biden when the state with the most votes for Trump  was 
``r max_percent_trump``. We then looked for the state that has the highest population of Black residents which happens to be 
``r max_race_pop_black``. We found this to be surprising since ``r max_race_pop_black``
was a red state in the 2020 election. We decided to challenge this discovery and calculate the voter turnout
in the state which ended up being ``r voter_turnout_ms``%. This rose a lot of hypotheses between our team such as voter suppression. 
We then decided to check the state with the highest white population and we discovered it was
``r max_race_pop_white``, which was insfact a red state as well. 

# Render table and describe pertinent observations found in it 
```{r summaryInformationTable, echo = FALSE, message =  FALSE}
```
This table reveals many insights to voter behavior. It shows total  individuals voted, 
total population, and total number registered from each race and gender. We can gather 
a lot of information on who is more likely to vote and what other affects their access 
to voting. 

# Describe the purpose of each chart, render each chart, and describe pertinent
# information found in each chart 

##chart 1
```{r chartOne, echo = FALSE, message =  FALSE}
```
This fasted histogram chart is intended to show the relationship between men and women 
who were registered and voted. The data from this graph allows us to see the 
difference in voter turnout based on gender. It shows that more women were registered 
to vote and were more likely to vote.  

##chart 2
```{r chartTwo, echo = FALSE, message =  FALSE}
```
This histogram chart was intended to show the relationship between the number 
of white individuals who registered and the number of white individuals who 
voted. The data from this graph then allows us to compare the differences 
between races in order to asses whether or not there is a significant disparity
between access to voting even after being registered between race groups.

##chart 3
```{r chartThree, echo = FALSE, message =  FALSE}
```
These pie charts were created with the intentions of visualizing the relationship
between the number of BIPOC individuals who registered to vote versus the
number of BIPOC individuals who voted within the U.S. 2016 Presidential 
Election. Similar to the other two charts, while this chart highlights a
different demographic, it will allow us to compare the statistics 
of voter registration between different genders and race groups to observe any
possible disparities. 
